Bootstrapping Multi-atlas Segmentation Using Multiple Automatically Generated Templates for the Segmentation of the Whole Hippocampus and Subfields

### 2014-01-03 paper.Rnw
Nice updated winterburn xval plots and montage

\begin{figure}
  \begin{centering}
    \includegraphics[width=\textwidth]{figure/MA-MAGeTBrain-Schematic}
  \end{centering}
  \caption{A schematic illustration of basic \multiatlas{} segmentation and
\mb{} segmentation. In multi-atlas segmentation, manual labels from atlas images
are warped (propagated) into subject space by applying the transformations
estimated from nonlinear image registration. The resulting candidate labels
from all atlas images are then fused to create a final segmentation. In \mb{}
segmentation, a template library is created by sampling (either randomly or
representatively) from the subject images. Atlas labels are propagated to all
template images and then to each subject image (including those used in the template
library). The candidate the labels for a subject are then fused into a final
segmentation.  \label{fig:MAGeT}}
\end{figure}

@@

\begin{figure}
  \begin{centering}
    \includegraphics[width=\textwidth]{figure/WinterburnSubfieldLOOCVSchematic}
  \end{centering}
  \caption{A schematic illustration of the experimental set up of the
  leave-one-out cross-validation (LOOCV) subfield experiment.}
  \label{fig:LOOCV-schematic}
\end{figure}

@@

d = ddply(na.omit(wa.xval.casted), c('region'), function (df) {
  icc = ICC(df[4:5])
  data.frame(
    icc  = icc$results$ICC[3],
    f    = icc$results$F[3],
    lower= icc$results$"lower bound"[3],
    upper= icc$results$"upper bound"[3],
  )
})

@@

@@

\begin{sidewaysfigure}
    \begin{centering}
      \includegraphics[width=\linewidth]{figure/winterburn-atlas-montage/figure.pdf}
    \end{centering}
    \caption{Detailed subfield segmentation results for a single subject. In
    the upper left corner is the original high-resolution Winterburn atlas manual
    subfield segmentation; in the upper right corner is the Winterburn atlas
    segmentation subsampled from 0.3mm- to 0.9mm-isotropic voxels; in the lower left
    corner is the \mb{} segmentation of the subsampled Winterburn atlas images; in the
    lower right corner is the \mb{} segmentation of a separately acquired T1 BRAVO
    image of the same subject. In each segmentation, slices from the left
    hemisphere are shown in Talairach-like ICBM152 space: the first row shows axial
    slices from inferior to superior; the second row shows sagittal slices from
    lateral to medial; the third row shows coronal slices from anterior to
    posterior.  \label{fig:subfield-montage}}
\end{sidewaysfigure}

@@

\begin{figure}
  \begin{centering}
    \includegraphics[width=6in]{figure/ADNI1_SNT_MB_montage/montage.pdf}
  \end{centering}
  \caption{\snt{} and \mb{} segmentations for 30 ADNI subjects (10
  subjects randomly selected from each disease category in the subject pool used
  in Experiment 1). Sagittal slices are shown for each unlabelled T1-weighted
  anatomical image. \snt{} labels appear in green, and \mb{} labels appear in
  blue. Noted are examples of common segmentation idiosyncrasies: 
  {\em (a)} over-estimation of hippocampal head and  
  {\em (b)} translated segmentation (seen in \snt{} segmentations only); 
  {\em (c)} under-estimation of hippocampal body and
  {\em (d)} improper inclusion of the vestigial hippocampal sulcus by \mb{}.}
  \label{fig:ADNI-segmentations}
\end{figure}

>>>>

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{figure/MA-MAGeTBrain-Schematic}
  \caption{A schematic illustration of basic \multiatlas{} segmentation and
\mb{} segmentation. In multi-atlas segmentation, manual labels from atlas images
are warped (propagated) into subject space by applying the transformations
estimated from nonlinear image registration. The resulting candidate labels
from all atlas images are then fused to create a final segmentation. In \mb{}
segmentation, a template library is created by sampling (either randomly or
representatively) from the subject images. Atlas labels are propagated to all
template images and then to each subject image (including those used in the template
library). The candidate the labels for a subject are then fused into a final
segmentation.  \label{fig:MAGeT}}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figure/WinterburnSubfieldLOOCVSchematic}
  \caption{A schematic illustration of the experimental set up of the
  leave-one-out cross-validation (LOOCV) subfield experiment.}
  \label{fig:LOOCV-schematic}
\end{figure}

@@

#d = ddply(na.omit(wa.xval.casted), c('region'), function (df) {
#  icc = ICC(df[4:5])
#  data.frame(
#    icc  = icc$results$ICC[3],
#    f    = icc$results$F[3],
#    lower= icc$results$"lower bound"[3],
#    upper= icc$results$"upper bound"[3],
#  )
#})

@@

\begin{figure}[h!]
<<fig:WAxval-kappa>>=
wa.n.sim = read.csv('data/winterburn.xval.similarity.csv')
wa.n.sim$label = as.factor(wa.n.sim$label)
levels(wa.n.sim$label) <-list(
  "CA1"=1, "CA1"=101, "Subiculum"=2,"Subiculum"=102, 
  "CA4/DG"=4,"CA4/DG"=104, "CA2/CA3"=5,"CA2/CA3"=105, 
  "SR/SL/SM"=6,"SR/SL/SM"=106)

# kappa vs. region
ggplot(wa.n.sim, aes(x=as.factor(label),y=k)) + 
  geom_boxplot(fill='lightblue') + 
  xlab("Subregion") + 
  ylab("Dice's Similarity Coefficient (DSC)")
@
  \caption{Similarity of \mb{} segmentation of subregions and resampled
  Winterburn atlas labels at $0.9mm^3$ voxel resolution over all folds of a
  leave-one-out cross validation.}
\end{figure}

\begin{figure}[h!]
<<fig:WAxval-Bland-Altman>>=
wa.n.vol = read.csv('data/winterburn.xval.volumes.csv')
winterburn = read.csv('data/winterburn.volumes.csv')

wa.n.vol.melt = melt(wa.n.vol, id.vars=c('atlases','subject','res'))
winterburn.melt = melt(winterburn, id.vars=c('subject','res'))
winterburn.cast = dcast(winterburn.melt, subject + variable ~ res)
wa.n.all.vols = merge(wa.n.vol.melt, winterburn.cast, by=c('subject','variable'))
#wa.n.vols.mean = ddply(wa.n.all.vols,c('subject','variable'),function (df) {
#  data.frame(
#    mb=mean(df$value),
#    X.3=mean(df$'0.3'),
#    X.9=mean(df$'0.9'))})

levels(wa.n.all.vols$variable) <-list(
  "CA1"="X1", "CA1"="X101", "Subiculum"="X2","Subiculum"="X102", 
  "CA4/DG"="X4","CA4/DG"="X104", "CA2/CA3"="X5","CA2/CA3"="X105", 
  "SR/SL/SM"="X6","SR/SL/SM"="X106")

wa.n.all.vols$diff = wa.n.all.vols$'0.9' - wa.n.all.vols$value
wa.n.all.vols$mean = (wa.n.all.vols$'0.9' + wa.n.all.vols$value)/2

wa.n.xval.bland.limits = ddply(wa.n.all.vols,c("variable"),function(df) {
  data.frame(y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T))
})
ggplot(wa.n.all.vols, aes(x=mean, y=diff)) + 
  facet_wrap(~variable, ncol=3, scales='free_x') + 
  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
  geom_point(size=1) +  
  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
             data=wa.n.xval.bland.limits, size=0.3, alpha=0.5) + 
  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
             data = wa.n.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
  xlab(expression('Mean of resampled and \\mb subregion volume ' (mm^3))) + 
  ylab(expression('Difference between manual and \\mb subregion volume ' (mm^3))) + 
  scale_color_discrete(name="") + 
  theme(legend.direction = "horizontal", legend.position = "bottom", 
        axis.text.y = element_text(angle = 90, hjust=0.5), 
        axis.text = element_text(size = 8))
@
  \caption{Bland-Altman plots across all folds of Winterburn atlas cross validation.}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=5in]{figure/winterburn-atlas-montage/figure.pdf}
    \caption{\textbf{Detailed subfield segmentation results for a single subject.} In
    the upper left corner is the original high-resolution Winterburn atlas manual
    subfield segmentation; in the upper right corner is the Winterburn atlas
    segmentation subsampled from 0.3mm- to 0.9mm-isotropic voxels; in the lower left
    corner is the \mb{} segmentation of the subsampled Winterburn atlas image
    from a single fold of the cross-validation. In each segmentation, slices from the left
    hemisphere are shown in Talairach-like ICBM152 space: the first row shows axial
    slices from inferior to superior; the second row shows sagittal slices from
    lateral to medial; the third row shows coronal slices from anterior to
    posterior. \label{fig:subfield-montage}}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=6in]{figure/ADNI1_SNT_MB_montage/montage.pdf}
  \caption{\snt{} and \mb{} segmentations for 30 ADNI subjects (10
  subjects randomly selected from each disease category in the subject pool used
  in Experiment 1). Sagittal slices are shown for each unlabelled T1-weighted
  anatomical image. \snt{} labels appear in green, and \mb{} labels appear in
  blue. Noted are examples of common segmentation idiosyncrasies: 
  {\em (a)} over-estimation of hippocampal head and  
  {\em (b)} translated segmentation (seen in \snt{} segmentations only); 
  {\em (c)} under-estimation of hippocampal body and
  {\em (d)} improper inclusion of the vestigial hippocampal sulcus by \mb{}.}
  \label{fig:ADNI-segmentations}
\end{figure}
